`r params$doc_title`

### 2023-03-13 scripts/MSc.Rmd
summary gt tables

For all scenarios (a. all instruments, b. common+strong, c. common+weak, d. rare+strong, e. rare+weak) we need to describe our results for a smaller set of sample sizes.

```{r}
df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method])
df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method], "_GenArch_", name_GenArch)

@@

    df_metrics_all_scenarios[[true_method]] <- df_metrics_all_scenarios[[true_method]] %>% 
      left_join(eval(as.symbol(df_name)) %>% 
                  rename_with(.fn = ~ paste0(.x, "_", name_GenArch), .cols = -"N"), by=c("N"="N"))
  }
}
# name list
names(df_metrics_all_scenarios) <- c("scenario_ALL_INSTR", "STRONG_COMMON", "STRONG_RARE", "WEAK_COMMON", "WEAK_RARE")
```

>>>>

For all scenarios (a. all instruments, b. common+strong, c. common+weak, d. rare+strong, e. rare+weak) we need to describe our results for a smaller set of sample sizes. We create a list that holds a dataframe for each reference method (1:5). Each dataframe contains results from all scenarios.

```{r}
# we have assigned dataframes in the loop above!
df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method])
df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method], "_GenArch_", name_GenArch)

@@

    df_metrics_all_scenarios[[true_method]] <- df_metrics_all_scenarios[[true_method]] %>% 
      left_join(eval(as.symbol(df_name)) %>% 
                  rename_with(.fn = ~ paste0(.x, "_", name_GenArch), .cols = -"N"), by=c("N"="N"))
  }
}

# name list
names(df_metrics_all_scenarios) <- paste0("ref_", selected_true_methods)
```

```{r gt-summary-tabs}
# https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/
library(gt)
df.temp.long <- list()
df.temp.wide <- list()

sample_size_subset <- c(50000, 100000, 200000, 300000)

for(ref.method in seq_along(df_metrics_all_scenarios)) {

  ## MEAN ## ---------------------------------------------------------------
  # later plot N=50,000 all methods and different scenarios

  ## CREATE LONG FORMAT
  df_metrics_all_scenarios[[ref.method]] %>% 
    filter(N %in% sample_size_subset) %>% 
    select(N, contains("_mean_") & 
             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
           ) %>% 
    #mutate(N = format(N, scientific=FALSE)) %>% 
    pivot_longer(cols=!N, names_to="Method", values_to="Mean")  %>% 
    separate(
      col="Method", into = c("Method", "Scenario"), sep = "_GSD_mean_"
    ) -> df.temp.long[[ref.method]]

  ## EMP SE ## ---------------------------------------------------------------
  df_metrics_all_scenarios[[ref.method]] %>% 
    filter(N %in% sample_size_subset) %>% 
    select(N, contains("_emp_se_") & 
             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
           ) %>% 
    #mutate(N = format(N, scientific=FALSE)) %>% 
    pivot_longer(cols=!N, names_to="Method", values_to="emp_se")  %>% 
    separate(
      col="Method", into = c("Method", "Scenario"), sep = "_GSD_emp_se_"
    )  -> temp.dat

  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))

  ## BIAS ## ---------------------------------------------------------------
  df_metrics_all_scenarios[[ref.method]] %>% 
    filter(N %in% sample_size_subset) %>% 
    select(N, contains("GSD_bias_") & 
             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
           ) %>% 
    #mutate(N = format(N, scientific=FALSE)) %>% 
    pivot_longer(cols=!N, names_to="Method", values_to="bias")  %>% 
    separate(
      col="Method", into = c("Method", "Scenario"), sep = "_GSD_bias_"
    )  -> temp.dat

  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))

  ## MSE ## ---------------------------------------------------------------
  df_metrics_all_scenarios[[ref.method]] %>% 
    filter(N %in% sample_size_subset) %>% 
    select(N, contains("_mse_") & 
             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
           ) %>% 
    #mutate(N = format(N, scientific=FALSE)) %>% 
    pivot_longer(cols=!N, names_to="Method", values_to="mse")  %>% 
    separate(
      col="Method", into = c("Method", "Scenario"), sep = "_GSD_mse_"
    )  -> temp.dat

  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))

  ## COVERAGE ## ---------------------------------------------------------------
  df_metrics_all_scenarios[[ref.method]] %>% 
    filter(N %in% sample_size_subset) %>% 
    select(N, contains("GSD_coverage_calc_") & 
             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
           ) %>% 
    #mutate(N = format(N, scientific=FALSE)) %>% 
    pivot_longer(cols=!N, names_to="Method", values_to="coverage")  %>% 
    separate(
      col="Method", into = c("Method", "Scenario"), sep = "_GSD_coverage_calc_"
    )  -> temp.dat

  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))

  ## TYPE1ERROR ## ---------------------------------------------------------------
  df_metrics_all_scenarios[[ref.method]] %>% 
    filter(N %in% sample_size_subset) %>% 
    select(N, contains("GSD_type1error_calc_") & 
             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
           ) %>% 
    #mutate(N = format(N, scientific=FALSE)) %>% 
    pivot_longer(cols=!N, names_to="Method", values_to="type1error")  %>% 
    separate(
      col="Method", into = c("Method", "Scenario"), sep = "_GSD_type1error_calc_"
    )  -> temp.dat

  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))

  #--
  ### CREATE WIDE FORMAT -------------------------------------------------------
  #--

  ## only one variable 
  #df.temp.long[[ref.method]] %>% 
  #  select(N, Method, Scenario, Mean) %>% 
  #  pivot_wider(names_from = "Scenario", values_from = "Mean", names_glue = "Mean_{Scenario}")

  ## several variables
  df.temp.long[[ref.method]] %>% 
    select(N, Method, Scenario, Mean, emp_se) %>% 
    pivot_wider(names_from = Scenario, values_from = c(Mean, emp_se),  
                #names_vary = "slowest",  # naming scheme
                names_glue = "{.value}_{Scenario}") -> df.temp.wide[[ref.method]]

# name list
names(df.temp.long) <- paste0("ref_", selected_true_methods)

### {gt} manual stub and grouping of rows  ###
### https://gt.rstudio.com/articles/intro-creating-gt-tables.html 
### https://themockup.blog/static/resources/gt-cookbook.html#introduction

#df.temp.wide[[ref.method]] %>% 
  #gt(groupname_col="N") %>% 
  #gt(rowname_col = "N") %>% 
  # tab_stubhead(label = "Sample Size") %>% 
  # tab_row_group(
  #   label = "Sample Size = 50,000", 
  #   rows = 1:10
  # ) %>% 
  # tab_row_group(
  #   label = "Sample Size = 100,000", 
  #   rows = 11:20
  # )

gt.tables.mean.se <- list()
gt.tables.performance2 <- list()

## loop to create tables
for (ref.method in seq_along(selected_true_methods)) {

  ## TABLE 01: MEAN +- SE
  df.temp.long[[ref.method]] %>% 
    select(N, Method, Scenario, Mean, emp_se) %>% 
    # first pivot wider
    pivot_wider(names_from = Scenario, values_from = c(Mean, emp_se),  
                  #names_vary = "slowest",  # naming scheme
                  names_glue = "{.value}_{Scenario}") %>% 
    mutate(N = paste0("Sample Size = ", format(N, scientific=FALSE))) %>% 
    gt(groupname_col="N") %>% 
    tab_header(
      title = md("**Mean estimate with corresponding empirical standard errors**"),
      subtitle = md(paste0("Reference estimate: ", selected_true_methods[ref.method]))
    ) %>% 
    tab_spanner(
      label = "Mean [\u00b1 Empirical Standard Error]", 
      columns = 3:7
    ) %>% 
    cols_label(
      Mean_ALL_INSTR = md("**Scenario 1**"),
      Mean_STRONG_COMMON = md("**Scenario 2**"),
      Mean_STRONG_RARE = md("**Scenario 3**"),
      Mean_WEAK_COMMON = md("**Scenario 4**"),
      Mean_WEAK_RARE = md("**Scenario 5**"),
    ) %>% 
    cols_merge(
      columns = c(Mean_ALL_INSTR, emp_se_ALL_INSTR),
      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
    ) %>% 
    cols_merge(
      columns = c(Mean_STRONG_COMMON, emp_se_STRONG_COMMON),
      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
    ) %>% 
    cols_merge(
      columns = c(Mean_STRONG_RARE, emp_se_STRONG_RARE),
      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
    ) %>% 
    cols_merge(
      columns = c(Mean_WEAK_COMMON, emp_se_WEAK_COMMON),
      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
    ) %>% 
    cols_merge(
      columns = c(Mean_WEAK_RARE, emp_se_WEAK_RARE),
      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
    ) %>% 
    fmt_number(columns = 3:12,decimals = 3, use_seps=TRUE) %>% 
    tab_source_note(
      source_note = md("*Scenario 1* corresponds to all genetic instruments being used; 
                       *Scenario 2* uses relatively strong and common variants;
                       *Scenario 3* uses relatively strong and rare variants; 
                       *Scenario 4* uses relatively weak and common variants;
                       *Scenario 5* uses relatively weak and rare variants")
    ) %>% 
    opt_stylize(style=1, color="gray") -> gt.tables.mean.se[[ref.method]]

  if (SAVE.files ==TRUE ){
    filename.gt.mean.se <- paste0("../output/gt-tables/", Sys.Date(), 
                                          "_mean_emp_se_performance_metrics_TRUE_EST_", 
                                          selected_true_methods[ref.method], 
                                          "_NSIM_", nsim)

    gt.tables.mean.se[[ref.method]] %>% 
      gtsave(paste0(filename.gt.mean.se, ".docx"))   # alternative .rtf, .tex, .png, .html
    gt.tables.mean.se[[ref.method]] %>% 
      gtsave(paste0(filename.gt.mean.se, ".html")) 
  }

  # TABLE 02: BIAS, MSE, COVERAGE, TYPE1ERROR
  df.temp.long[[ref.method]] %>%
    select(N, Method, Scenario, bias, mse, coverage, type1error) %>% 
    # first pivot wider
    pivot_wider(names_from = Scenario, values_from = c(bias, mse, coverage, type1error),  
                  #names_vary = "slowest",  # naming scheme
                  names_glue = "{.value}_{Scenario}") %>% 
    mutate(N = paste0("Sample Size = ", format(N, scientific=FALSE))) %>% 
    gt(groupname_col="N") %>% 
    tab_header(
      title = md("**Performance measures**"),
      subtitle = md(paste0("Reference estimate: ", selected_true_methods[ref.method]))
    ) %>% 
    tab_spanner(
      label = "Bias", 
      columns = 3:7
    ) %>% 
    tab_spanner(
      label = "MSE", 
      columns = 8:12
    ) %>% 
    tab_spanner(
      label = "Coverage", 
      columns = 13:17
    ) %>% 
    tab_spanner(
      label = "Type 1 Error", 
      columns = 18:22
    ) %>% 
    cols_label(
      bias_ALL_INSTR = md("**Scenario 1**"),
      bias_STRONG_COMMON = md("**Scenario 2**"),
      bias_STRONG_RARE = md("**Scenario 3**"),
      bias_WEAK_COMMON = md("**Scenario 4**"),
      bias_WEAK_RARE = md("**Scenario 5**"),
    ) %>% 
    cols_label(
      mse_ALL_INSTR = md("**Scenario 1**"),
      mse_STRONG_COMMON = md("**Scenario 2**"),
      mse_STRONG_RARE = md("**Scenario 3**"),
      mse_WEAK_COMMON = md("**Scenario 4**"),
      mse_WEAK_RARE = md("**Scenario 5**"),
    ) %>% 
    cols_label(
      coverage_ALL_INSTR = md("**Scenario 1**"),
      coverage_STRONG_COMMON = md("**Scenario 2**"),
      coverage_STRONG_RARE = md("**Scenario 3**"),
      coverage_WEAK_COMMON = md("**Scenario 4**"),
      coverage_WEAK_RARE = md("**Scenario 5**"),
    ) %>% 
    cols_label(
      type1error_ALL_INSTR = md("**Scenario 1**"),
      type1error_STRONG_COMMON = md("**Scenario 2**"),
      type1error_STRONG_RARE = md("**Scenario 3**"),
      type1error_WEAK_COMMON = md("**Scenario 4**"),
      type1error_WEAK_RARE = md("**Scenario 5**"),
    ) %>% 
    fmt_number(columns = 3:22,decimals = 3, use_seps=TRUE) %>% 
    tab_source_note(
      source_note = md("*Scenario 1* corresponds to all genetic instruments being used; 
                       *Scenario 2* uses relatively strong and common variants;
                       *Scenario 3* uses relatively strong and rare variants; 
                       *Scenario 4* uses relatively weak and common variants;
                       *Scenario 5* uses relatively weak and rare variants")
    ) %>% 
    opt_stylize(style=1, color="gray") -> gt.tables.performance2[[ref.method]]

    if (SAVE.files ==TRUE ){
      filename.gt.performance2 <- paste0("../output/gt-tables/", Sys.Date(), 
                                            "_performance2_metrics_TRUE_EST_", 
                                            selected_true_methods[ref.method], 
                                            "_NSIM_", nsim)

      gt.tables.performance2[[ref.method]] %>% 
        gtsave(paste0(filename.gt.performance2, ".docx"))   # alternative .rtf, .tex, .png, .html
      gt.tables.performance2[[ref.method]] %>% 
        gtsave(paste0(filename.gt.performance2, ".html")) 
    }

  ## --
  # CREATE PLOTS -------------------------------------------------------------
  ## --

  df.temp.long[[ref.method]] %>%
    mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
    ggplot(aes(x=Method, y=mse, fill=Scenario )) +
    geom_bar(stat="identity", position="dodge") +
    facet_wrap(~N) + 
    coord_flip()+
    ggthemes::theme_few()+
    ggthemes::scale_fill_few() -> dodged.barplot.mse
